---
title: "strings and factors"
output: github_document
date: "2025-10-14"
---

```{r}
library(tidyverse)
library(rvest)
```

## Mostly use string vectors
note`str_detect` is case sensitive
```{r}
string_vec = c("my", "name", "is", "chelsey")

str_detect(string_vec, "chelsey")
str_detect(string_vec, "e")

str_replace(string_vec, "chelsey", "Chelsey")
str_replace(string_vec, "e", "E")

str_replace(string_vec, "e", "")
str_remove(string_vec, "e")
```

note: `str_detect(string_vec, "i think$ ")` aka looking for "i think" at the end of sentences
note: `str_detect(string_vec, "^i think")` aka looking for "i think" at beginning of sentences
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")
str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$ ")

str_remove(string_vec, "i think$")

```

`[Pp]` looking for anything w/in brackets
```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )


str_detect(string_vec, "[Pp]umpkin")

```


Let's get a bit more complicated
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-z]")
```

note: `.` represents that anything can be in that place
note: `"7\\.1" aka looking for "7.1" becasue `\\` let's you know you are looking for a special character
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

 str_detect(string_vec, "7\\.1") 
```

## Factors

```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex
```

```{r}

```


