---
title: "strings and factors"
output: github_document
date: "2025-10-14"
---

```{r}
library(tidyverse)
library(rvest)
library(p8105.datasets)
```

## Mostly use string vectors
note`str_detect` is case sensitive
```{r}
string_vec = c("my", "name", "is", "chelsey")

str_detect(string_vec, "chelsey")
str_detect(string_vec, "e")

str_replace(string_vec, "chelsey", "Chelsey")
str_replace(string_vec, "e", "E")

str_replace(string_vec, "e", "")
str_remove(string_vec, "e")
```

note: `str_detect(string_vec, "i think$ ")` aka looking for "i think" at the end of sentences
note: `str_detect(string_vec, "^i think")` aka looking for "i think" at beginning of sentences
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")
str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$ ")

str_remove(string_vec, "i think$")

```

`[Pp]` looking for anything w/in brackets
```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )


str_detect(string_vec, "[Pp]umpkin")

```


Let's get a bit more complicated
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-z]")
```

note: `.` represents that anything can be in that place
note: `"7\\.1" aka looking for "7.1" becasue `\\` let's you know you are looking for a special character
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

 str_detect(string_vec, "7\\.1") 
```

## Factors

```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex
```

```{r}
as.numeric(vec_sex)
```

```{r}
vec_sex = fct_inorder(vec_sex)
vec_sex
```

## Revisit NSDOH

import NSDUH data from the web
```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)
```

this is an "easy case
note:'slice' function allows you to take the first row out
```{r}
nsduh_df =
  drug_use_html %>% 
  html_table() %>% 
  first() %>% 
  slice(-1)
```

now do the tidying....
note: `filter(
    !(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West"))` means removes where State variable is not in the collection of regions I stated above
```{r}
marj_df = 
  nsduh_df %>% 
  select(-contains("P Value")) %>% 
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
  ) %>% 
  separate(age_year, into = c("age", "year"), sep = "\\(") %>% 
  mutate(
    year = str_remove(year, "\\)"),
    percent = str_remove(percent, "[a-c]$"),
    percent = as.numeric(percent)
  ) %>% 
  filter(
    !(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West"))
  )
```

Let's make a quick plot
note: ` mutate(State = fct_reorder(State, percent))` means you are reordering State variable by percent variable
```{r}
marj_df %>% 
  filter(age == "12-17") %>%
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust=1))
```

## Restaurant and Inspection
```{r}
data("rest_inspec")
```


```{r}
rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarise(n = n()) %>% 
  pivot_wider(
    names_from = grade,
    values_from = n
  )
```

update the dataset
```{r}
rest_inspec =
  rest_inspec %>% 
  filter(
    grade %in% c("A", "B", "C"),
         boro != 0)
```

look for pizza places
```{r}
rest_inspec %>% 
  filter(str_detect(dba, "Pizza"))
```

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  group_by(boro, grade) %>% 
  summarise(n = n()) %>% 
  pivot_wider(
    names_from = grade,
    values_from = n
  )
```

let's make this table a plot instead
```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(boro = fct_infreq(boro)) %>% 
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```


```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(boro = fct_infreq(boro),
         boro = fct_recode(boro, "The City" = "Manhattan")) %>% 
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```